---
title: "USE_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{USE_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we will go trough the USE package. 

```{r setup, message=FALSE}
Sys.setlocale("LC_ALL", "English")
library(USE)
library(raster)
library(virtualspecies)
library(RStoolbox)
library(data.table)
library(sf)
myCRS<-"+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
```

Download bioclimatic variables from WorldClim
```{r, eval=TRUE, message=FALSE}
Worldclim<-raster::getData('worldclim', var='bio', res=10) 
```
Crop the bioclimatic variables to the European extent
```{r, eval=TRUE}
envData<-crop(Worldclim, extent(-12, 25, 36, 60))
envData
```

Create virtual species. For details about the methodology used, see the  [vignette](http://borisleroy.com/files/virtualspecies-tutorial.html#introduction) of the **virtualspecies** R package
```{r, eval=TRUE, message=FALSE}
#create virtual species
myRandNum=sample(1:19,size=5, replace = FALSE)
random.sp <- virtualspecies::generateRandomSp(envData[[myRandNum]], 
                                              convert.to.PA = FALSE, 
                                              species.type = "additive",
                                              realistic.sp = TRUE, 
                                              plot = FALSE)
#reclassify suitability raster using a probability conversion rule
new.pres<-convertToPA(x=random.sp, 
                      beta=0.55,
                      alpha = -0.05, plot = FALSE)
#Sample true occurrences
presence.points <- sampleOccurrences(new.pres,
                                     n = 300, # The number of points to sample
                                     type = "presence-absence",
                                     sample.prevalence = 0.6,
                                     detection.probability = 1,
                                     correct.by.suitability = TRUE,
                                     plot = TRUE)  
```

Get a presence only dataset.
```{r, eval=TRUE}
myPres=presence.points$sample.points[which(presence.points$sample.points$Observed==1), c( "x", "y",  "Observed")]
coordinates(myPres)<-~x+y
crs(myPres)<-myCRS
myPres
```

Find optimal resolution of the sampling grid. First, made a PCA on the whole raster stack and then use the PC-scores of the first two axes to create a new spatial object. 
```{r, eval=TRUE}
rpc<-rasterPCA(envData,spca = TRUE)
dt <- na.omit(data.table(as.data.frame(rpc$map[[c("PC1", "PC2")]], xy = TRUE)))
dt=st_as_sf(dt, coords = c("PC1", "PC2"))
```

```{r, eval=FALSE, echo=TRUE}
myRes=USE::optimRes(sdf=dt,
                    grid.res=c(1:10),
                    perc.thr = 10,
                    showOpt = TRUE)
```
```{r, eval=TRUE, echo=FALSE, message=FALSE, results="hide"}
myRes=list()
myRes$Opt_res
myRes$Opt_res=6
```
```{r, eval=TRUE}
myRes$Opt_res
```

Perform uniform sampling of the environmental space
```{r, eval=TRUE, message=FALSE}
myObs=USE::uniformSampling(sdf=dt, 
                           grid.res=myRes$Opt_res,
                           n.tr = 5,
                           sub.ts = TRUE,
                           n.ts = 2,
                           plot_proc = FALSE)
```
```{r, eval=TRUE}
str(myObs)
#background points for the training dataset
myObs$Bkg.tr
#background points for the testing dataset
myObs$Bkg.ts
```

Uniformly sampling background points in the environmental space, accounting for the species presence points. 
```{r, eval=TRUE, message=FALSE}
myGrid.psAbs <- USE::paSampling(env.rast=envData,
                                pres=myPres,
                                thres=0.75,
                                H=NULL,
                                grid.res=as.numeric(myRes$Opt_res),
                                n.tr = 5,
                                prev=0.3,
                                sub.ts=TRUE,
                                n.ts=5,
                                plot_proc=FALSE,
                                verbose=FALSE)
```

```{r, eval=TRUE}
str(myGrid.psAbs)
#background points for the training dataset
myGrid.psAbs$Bkg.tr
#background points for the testing dataset
myGrid.psAbs$Bkg.ts
```

